---
title: "Word order constraints on event-internal modifiers - Experiment 2: Forced Choice study on Thematic Integration"
author: "Tibor Kiss"
date: "10.11.2021"
output:
  github_document:
    keep_html: TRUE
---

### Analysis of experimental study on thematic integration

This document provides the analysis for the experimental study on thematic integration used in Kiss, Pieper, Börner (2021), final version, including the random slopes on the interaction of main effects.

#### Read in required libraries

```{r libraries, warning=FALSE, message=FALSE}
library(lme4)
library(tidyverse)
library(emmeans)

show(version)
```

#### Read in data set


```{r data set, cache = TRUE}
exp_int_agentivity.fc.test.data <- 
  read.csv("../Data/ForcedChoice_test.csv", fileEncoding = "UTF-8")

exp_int_agentivity.fc.test.data$ANSWER <- 
  factor(exp_int_agentivity.fc.test.data$ANSWER)

exp_int_agentivity.fc.test.data$INTEGRATION <- 
  factor(exp_int_agentivity.fc.test.data$INTEGRATION)

exp_int_agentivity.fc.test.data$ADVERBIAL_TYPE <- 
  factor(exp_int_agentivity.fc.test.data$ADVERBIAL_TYPE)

## more illustrative variable names
exp_int_agentivity.fc.test.data$subjects <- factor(exp_int_agentivity.fc.test.data$workerId)
exp_int_agentivity.fc.test.data$items <- factor(exp_int_agentivity.fc.test.data$ENCODING)

```

The following table shows an equal distribution of test items to subjects (number of accepted subjects: `r length(table(exp_int_agentivity.fc.test.data$workerId))`).

```{r, echo = FALSE}
table(exp_int_agentivity.fc.test.data$workerId)
```

#### Empirical distribution of choices


```{r empirical distribution, cache = TRUE, warning = FALSE, message = FALSE, fig.width = 8}

thematic_int.labels <- c("No Thematic Integration", "Thematic Integration")
names(thematic_int.labels) <- c("no", "yes")

exp_int_agentivity.fc.summary <- 
  exp_int_agentivity.fc.test.data %>%
  group_by(ADVERBIAL_TYPE, INTEGRATION, ANSWER) %>%
  summarise(count = n()) %>%
  as.data.frame()

ggplot(exp_int_agentivity.fc.summary,
       aes(x = ANSWER, y = count, fill = ADVERBIAL_TYPE)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Choice", y = "Frequency of Choice") + 
  facet_wrap(~INTEGRATION, labeller = labeller(INTEGRATION = thematic_int.labels)) +
  scale_fill_manual("ADVERBIAL_TYPE", values = c("COM(S)" = "black",
                                                 "INSTR" = "grey"),
                    labels = c("comitative (subj.)", "instrumental")) +
  guides(fill=guide_legend(title="Semantic type\nof adverbial")) 

```

The graphical representation of the empirical distribution of choices suggests an interaction between the main effects, as is witnessed by the inversion of the distribution of `COM(S)` under `Thematic Integration == "yes"`. 

#### Random Slope Model with interaction

Please notice that the random structure does not assume intercepts, because we want to obtain information on by-subject variance (i.e. workerId) for each condition.

```{r random slope model, cache = TRUE}

exp_int_agentivity.fc.test.data$INTEGRATION <-
  relevel(exp_int_agentivity.fc.test.data$INTEGRATION, ref = "no")
exp_int_agentivity.fc.test.data$ADVERBIAL_TYPE <-
  relevel(exp_int_agentivity.fc.test.data$ADVERBIAL_TYPE, ref = "INSTR")


exp_int_agentivity.fc.glmm <- 
  glmer(formula = ANSWER ~ ADVERBIAL_TYPE * INTEGRATION + 
          (0 + ADVERBIAL_TYPE * INTEGRATION | subjects) , 
        data = exp_int_agentivity.fc.test.data, family = binomial())

options(width = 300)
summary(exp_int_agentivity.fc.glmm)
```

We'll have to consider possible spurious convergence warnings, and apply tests from https://rstudio-pubs-static.s3.amazonaws.com/33653_57fc7b8e5d484c909b615d8633c01d51.html, i.e. Ben Bolker's troubleshooting page for spurious convergence errors.

```{r singularity, cache = TRUE}

tt <- getME(exp_int_agentivity.fc.glmm,"theta")
ll <- getME(exp_int_agentivity.fc.glmm,"lower")
```

The comparatively high value of `r round(min(tt[ll==0]), 3)` suggests that singularity is not an issue here. Restarting solves the problem here. 

```{r restarting, cache = TRUE}
ss <- 
  getME(exp_int_agentivity.fc.glmm,c("theta","fixef"))
exp_int_agentivity.fc.glmm <- 
  update(exp_int_agentivity.fc.glmm,
         start=ss,
         control=glmerControl(optCtrl=list(maxfun=2e4)))

options(width = 300)
summary(exp_int_agentivity.fc.glmm)

```

It is an interesting fact that the by-subject variation for `COM(S):INTEGRATION yes` is much lower than for the other conditions, which appears to confirm the preference for PP > OBJ in this condition.

#### Model predictions


```{r Model predictions, fig.width = 8}


predictions <- 
  emmeans(exp_int_agentivity.fc.glmm, pairwise~ADVERBIAL_TYPE * INTEGRATION, 
          type = "response") 

## also response useful, now switched to predictor for effects
## if lp is used, y = emmean, if response is used y = prob

predictions.emm <- data.frame(predictions$emmeans)
 

predictions.emm

ggplot(predictions.emm, aes(x = ADVERBIAL_TYPE, y = prob)) + 
  geom_bar(aes(), stat = "identity", fill = "grey", width = 0.3) +
  facet_wrap(~INTEGRATION, labeller = labeller(INTEGRATION = thematic_int.labels)) +
  labs(x = "Adverbial Types", y = "Pr(PP>OBJ)")
```


