---
title: "Word order constraints on event-internal modifiers - Experiment 2: Forced Choice study on Thematic Integration"
author: "Tibor Kiss"
date: "10.11.2021"
output:
  github_document:
    keep_html: TRUE
---

### Analysis of experimental study on thematic integration

This document provides the analysis for the experimental study on thematic integration used in Kiss, Pieper, Börner (2021), final version, including the random slopes on the interaction of main effects.

#### Read in required libraries

```{r libraries, warning=FALSE, message=FALSE}
library(lme4)
library(tidyverse)
library(emmeans)

show(version)
```

#### Read in data set


```{r data set, cache = TRUE}
exp_int_agentivity.fc.test.data <- 
  read.csv("../Data/ForcedChoice_test.csv", fileEncoding = "UTF-8")

exp_int_agentivity.fc.test.data$ANSWER <- 
  factor(exp_int_agentivity.fc.test.data$ANSWER)

exp_int_agentivity.fc.test.data$INTEGRATION <- 
  factor(exp_int_agentivity.fc.test.data$INTEGRATION)

exp_int_agentivity.fc.test.data$ADVERBIAL_TYPE <- 
  factor(exp_int_agentivity.fc.test.data$ADVERBIAL_TYPE)

## more illustrative variable names
exp_int_agentivity.fc.test.data$subjects <- factor(exp_int_agentivity.fc.test.data$workerId)
exp_int_agentivity.fc.test.data$items <- factor(exp_int_agentivity.fc.test.data$ENCODING)

```

The following table shows an equal distribution of test items to subjects (number of accepted subjects: `r length(table(exp_int_agentivity.fc.test.data$workerId))`).

```{r, echo = FALSE}
table(exp_int_agentivity.fc.test.data$workerId)
```

#### Empirical distribution of choices


```{r empirical distribution, cache = TRUE, warning = FALSE, message = FALSE, fig.width = 8}

thematic_int.labels <- c("No Thematic Integration", "Thematic Integration")
names(thematic_int.labels) <- c("no", "yes")

exp_int_agentivity.fc.summary <- 
  exp_int_agentivity.fc.test.data %>%
  group_by(ADVERBIAL_TYPE, INTEGRATION, ANSWER) %>%
  summarise(count = n()) %>%
  as.data.frame()

ggplot(exp_int_agentivity.fc.summary,
       aes(x = ANSWER, y = count, fill = ADVERBIAL_TYPE)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Choice", y = "Frequency of Choice") + 
  facet_wrap(~INTEGRATION, labeller = labeller(INTEGRATION = thematic_int.labels)) +
  scale_fill_manual("ADVERBIAL_TYPE", values = c("COM(S)" = "black",
                                                 "INSTR" = "grey"),
                    labels = c("comitative (subj.)", "instrumental")) +
  guides(fill=guide_legend(title="Semantic type\nof adverbial")) 

```

The graphical representation of the empirical distribution of choices suggests an interaction between the main effects, as is witnessed by the inversion of the distribution of `COM(S)` under `Thematic Integration == "yes"`. 

#### Random Slope Model with interaction

Please notice that the random structure does not assume intercepts, because we want to obtain information on by-subject variance (i.e. workerId) for each condition.

```{r random slope model, cache = TRUE}

exp_int_agentivity.fc.test.data$INTEGRATION <-
  relevel(exp_int_agentivity.fc.test.data$INTEGRATION, ref = "no")
exp_int_agentivity.fc.test.data$ADVERBIAL_TYPE <-
  relevel(exp_int_agentivity.fc.test.data$ADVERBIAL_TYPE, ref = "INSTR")


exp_int_agentivity.fc.glmm <- 
  glmer(formula = ANSWER ~ ADVERBIAL_TYPE * INTEGRATION + 
          (0 + ADVERBIAL_TYPE * INTEGRATION | subjects) , 
        data = exp_int_agentivity.fc.test.data, family = binomial())

options(width = 300)
summary(exp_int_agentivity.fc.glmm)
```

We'll have to consider possible spurious convergence warnings, and apply tests from https://rstudio-pubs-static.s3.amazonaws.com/33653_57fc7b8e5d484c909b615d8633c01d51.html, i.e. Ben Bolker's troubleshooting page for spurious convergence errors.

```{r singularity, cache = TRUE}

tt <- getME(exp_int_agentivity.fc.glmm,"theta")
ll <- getME(exp_int_agentivity.fc.glmm,"lower")
```

The comparatively high value of `r round(min(tt[ll==0]), 3)` suggests that singularity is not an issue here. Restarting solves the problem here. 

```{r restarting, cache = TRUE}
ss <- 
  getME(exp_int_agentivity.fc.glmm,c("theta","fixef"))
exp_int_agentivity.fc.glmm <- 
  update(exp_int_agentivity.fc.glmm,
         start=ss,
         control=glmerControl(optCtrl=list(maxfun=2e4)))

options(width = 300)
summary(exp_int_agentivity.fc.glmm)

```

If we compare the two alternatives for INTEGRATION, we notice two things:

* The by-subject variation for `INTEGRATION == "no"` is similar for both `ADVERBIAL_TYPEs`, and there is a positive correlation between the random effects for `INSTR` (random intecept) and `COM(S)` (random slope), which means that subjects may have individual preferences, but these are the same for both `ADVERBIAL_TYPEs`. 
* In contrast, the by-subject variation for `INTEGRATION == "yes"` differs sharply: For `INSTR` it is comparable with the values for `INTEGRATION == "no"`, but for `COM(S)` it is much smaller. In addition, there is a negative correlation between the random slopes for  `ADVERBIAL_TYPE`, so that lower probabilities for chosing `PP > OBJ` given `INSTR` correspond to higher probabilities for chosing `PP > OBJ` given `COM(S)`.

The model of the experiment thus suggests that there is considerable variance among speakers in case of instrumentals or no thematic integration, and much less variance in case of thematic integration with comitatives. (The correlations are illustrated below.)

#### Model predictions


```{r Model predictions, fig.width = 8}


predictions <- 
  emmeans(exp_int_agentivity.fc.glmm, pairwise~ADVERBIAL_TYPE * INTEGRATION, 
          type = "response") 

## also response useful, now switched to predictor for effects
## if lp is used, y = emmean, if response is used y = prob

predictions.emm <- data.frame(predictions$emmeans)
 

predictions.emm

ggplot(predictions.emm, aes(x = ADVERBIAL_TYPE, y = prob)) + 
  geom_bar(aes(), stat = "identity", fill = "grey", width = 0.3) +
  facet_wrap(~INTEGRATION, labeller = labeller(INTEGRATION = thematic_int.labels)) +
  labs(x = "Adverbial Types", y = "Pr(PP>OBJ)")
```

<<<<<<< Updated upstream

=======
#### Plotting correlations for random effects

In the following, we plot the correlations between random effects for `INTEGRATION == "no"` and `INTEGRATION == "yes"`, as individual odds ratios. (I am not sure whether this is actually best practice, but it is illustrative.)

```{r}

ranef.df <- data.frame(ranef(exp_int_agentivity.fc.glmm)$subjects)
ranef.df$participants <- rownames(ranef.df)
colnames(ranef.df)[1:4] <- c("INSTR", "COM_S", "INT_YES_INSTR", "INT_YES_COM_S")

ranef.df <- ranef.df %>%
  mutate(INSTR = INSTR + exp_int_agentivity.fc.glmm@beta[1]) %>%
  mutate(COM_S = COM_S + exp_int_agentivity.fc.glmm@beta[2]) %>%
  mutate(INT_YES_INSTR = INT_YES_INSTR + exp_int_agentivity.fc.glmm@beta[3]) %>%
  mutate(INT_YES_COM_S = INT_YES_COM_S + exp_int_agentivity.fc.glmm@beta[4]) %>%
  arrange(INT_YES_COM_S)

ggplot(ranef.df, aes(x = INT_YES_COM_S, y = INT_YES_INSTR)) +
  geom_point(aes(color = participants), show.legend = FALSE) +
  labs(x = "PP > OBJ given COM(S)", 
       y = "PP > OBJ given INSTR", 
       subtitle = "Negative correlation of random slopes for ADVERBIAL_TYPE given INTEGRATION == yes")

ggplot(ranef.df, aes(x = COM_S, y = INSTR)) +
  geom_point(aes(color = participants), show.legend = FALSE) +
  labs(x = "PP > OBJ given COM(S)", 
       y = "PP > OBJ given INSTR", 
       subtitle = "Positive correlation of random slopes for ADVERBIAL_TYPE given INTEGRATION == no")

```

We notice that not only do the correlations between the random effects for `ADVERBIAL_TYPE` differ, but also that the by-subject variability for `COM(S)` given `INTEGRATION == "yes"` is much smaller. 
>>>>>>> Stashed changes
