---
title: "Word order constraints on event-internal modifiers - Experiment 1: Likert Scale study on Anaphoricity"
author: "Tibor Kiss"
date: "10.01.2022"
output:
  github_document:
    keep_html: TRUE
---

```{r libraries, warning=FALSE, message=FALSE}
library(ordinal)
library(tidyverse)
library(emmeans)

show(version)
```

#### Introduction

This document describes the necessary analysis of Experiment 1 (LS) for event-internal modifiers (second stage). We have chosen a model with a rather complex random structure, which however, similar to the model for the forced choice study on thematic integration, reveals differences in by-subject variability.  


#### Read in data set

```{r data preparation}
exp_anaphoricity_ls.test.data <- 
  read.csv("../Data/LikertSkala_test.csv", fileEncoding = "UTF-8", stringsAsFactors = TRUE)

exp_anaphoricity_ls.test.data$CONDITION_NO <- factor(exp_anaphoricity_ls.test.data$CONDITION_NO)

exp_anaphoricity_ls.test.data$FCT_ANSWER <- 
  factor(exp_anaphoricity_ls.test.data$ANSWER, ordered = TRUE)

exp_anaphoricity_ls.test.data$ENCODING <- 
  factor(as.character(exp_anaphoricity_ls.test.data$ENCODING))

exp_anaphoricity_ls.test.data$KEY_CONDITION = relevel(exp_anaphoricity_ls.test.data$KEY_CONDITION, "PP>OBJ") ## CONDITION_NO 1

## more illustrative variable names
exp_anaphoricity_ls.test.data$POSITION <- factor(exp_anaphoricity_ls.test.data$KEY_CONDITION)
exp_anaphoricity_ls.test.data$subjects <- factor(exp_anaphoricity_ls.test.data$workerId)
exp_anaphoricity_ls.test.data$items <- factor(exp_anaphoricity_ls.test.data$ENCODING)

```

#### Empirical distribution of judgments

Number of accepted participants: `r length(unique(exp_anaphoricity_ls.test.data$workerId))`.

```{r distribution of items to participants, warning = FALSE, message = FALSE}
exp_anaphoricity_ls.dist.summary <- exp_anaphoricity_ls.test.data %>% 
  select(workerId, CONDITION_NO) %>% table() %>% as.data.frame()
exp_anaphoricity_ls.dist.summary %>% filter(Freq < 18)
```

Each participant has seen 18 test items per condition, since no participant has seen less than 18 test items per condition (8->). 

```{r distribution of judgments, fig.width = 10, message = FALSE, warning = FALSE}

exp_anaphoricity_ls.test.summary <- 
  exp_anaphoricity_ls.test.data %>% 
  group_by(POSITION, ADVERBIAL_TYPE, ANSWER) %>% 
  summarise(count = n())

ggplot(exp_anaphoricity_ls.test.summary, aes(x = ANSWER, y = count)) +
  geom_bar(stat = "identity", position = "dodge", show.legend = FALSE) +
  facet_wrap(POSITION~ADVERBIAL_TYPE) + 
  labs(x = "Ratings", y = "Frequency of Ratings")

```


### Random slope model for items and subjects without interaction

This model does not make use of interaction of the fixed effects but assumes an interaction for the random effects so that random effects for each condition (3 x 2) are modelled for subjects and thus can be compared. (Items of course do not vary w.r.t. `ADVERBIAL_TYPE`.)

#### Definition of model

```{r model, cache = TRUE, warning=FALSE}

exp_anaphoricity_ls.test.data$ADVERBIAL_TYPE <-
  relevel(factor(exp_anaphoricity_ls.test.data$ADVERBIAL_TYPE), ref = "INSTR")
exp_anaphoricity_ls.test.data$POSITION <-
  relevel(factor(exp_anaphoricity_ls.test.data$POSITION), ref = "PP>OBJ")

exp_anaphoricity_ls.clmm <- 
  clmm(FCT_ANSWER  ~ ADVERBIAL_TYPE + POSITION + 
         (0 + ADVERBIAL_TYPE * POSITION | subjects) +
         (0 + POSITION | items), 
       exp_anaphoricity_ls.test.data)
options(width = 300)
summary(exp_anaphoricity_ls.clmm)

exp_anaphoricity_ls.clmm2 <- 
  clmm(FCT_ANSWER  ~ ADVERBIAL_TYPE * POSITION + 
         (0 + ADVERBIAL_TYPE * POSITION| subjects) +
         (0 + POSITION | items), 
       exp_anaphoricity_ls.test.data)

summary(exp_anaphoricity_ls.clmm2)

```

A comparison of the two models shows that they are not significantly different. We will thus keep the model with interaction in the random effects but no interaction in the fixed effects:


```{r comparison of models}

comp <- anova(exp_anaphoricity_ls.clmm, exp_anaphoricity_ls.clmm2)

show(comp)
```

The random structure of the model reveals that the by-subject variability is reduced under `OBJ > PP`.


#### Effects in the model w/o interaction

Since we have chosen INSTR as reference value, the values for ILOC and COM(O) both show significance, but it should be kept in mind that they are practically identical. The general effect of `POSITION == OBJ > PP` is `r round(exp(exp_anaphoricity_ls.clmm$beta[3]), 3)`. The effects for COM(O) and ILOC in `POSITION == PP > OBJ` are `r round(exp(exp_anaphoricity_ls.clmm$beta[1]), 3)`, and `r round(exp(exp_anaphoricity_ls.clmm$beta[2]), 3)`, respectively.  


#### Model predictions 

```{r graphical presentation of model without interaction, fig.width = 10}

pred <-function(eta, theta, cat = 1:(length(theta)+1), inv.link= plogis){
  Theta <- c(-1e3, theta, 1e3)
  sapply(cat, function(j) inv.link(Theta[j+1] - eta) - inv.link(Theta[j] - eta))
}

exp_anaphoricity_ls.mat <- 
  data.frame(
    adv = rep(c(0, exp_anaphoricity_ls.clmm$beta[1], 
                exp_anaphoricity_ls.clmm$beta[2]), 2),
    cond = c(rep(0, 3), rep(exp_anaphoricity_ls.clmm$beta[3], 3))
  )

pred.mat <- pred(eta=rowSums(exp_anaphoricity_ls.mat), 
                 theta=exp_anaphoricity_ls.clmm$Theta)

exp_anaphoricity_ls.mat$condition <- 
  c(rep("PP>OBJ", 3), rep("OBJ>PP", 3))
exp_anaphoricity_ls.mat$adv_type <- 
  rep(c("INSTR", "COM(O)", "ILOC"), 2)

exp_anaphoricity_ls.pred <- 
  cbind(exp_anaphoricity_ls.mat, pred.mat)

exp_anaphoricity_ls.pred.long <- 
  exp_anaphoricity_ls.pred %>%
  gather("1":"5", key = "ANSWER", value = "rating")

exp_anaphoricity_ls.pred.long$condition <- 
  relevel(factor(exp_anaphoricity_ls.pred.long$condition), ref = "PP>OBJ")

ggplot(exp_anaphoricity_ls.pred.long, aes(x = ANSWER, y = rating)) + 
  geom_bar(stat = "identity", show.legend = FALSE) + 
  facet_wrap(condition ~ adv_type) + 
  labs(x = "Ratings", y = "Pr(Ratings)")


```


#### Test items in disfavored positions with high ratings 

```{r warning=FALSE, message = FALSE}

exp_anaphoricity_ls.rating.summary <- 
  subset(exp_anaphoricity_ls.test.data, 
         ADVERBIAL_TYPE == "COM(O)" & POSITION == "PP>OBJ") %>% 
  group_by(ENCODING, ANSWER) %>% 
  summarise(count = n()) %>%
  as.data.frame()

ggplot(exp_anaphoricity_ls.rating.summary, aes(x = ANSWER, y = count)) +
  geom_bar(stat = "identity") +
  facet_wrap(~ENCODING) +
  labs(x = "Rating", y = "Frequency of rating")
```

```{r}
read.csv("../Data/stimuli_test_basic.csv", fileEncoding = "UTF-8") %>% 
  filter(ADVERBIAL_TYPE == "COM(O)" & KEY_CONDITION == "PP>OBJ") %>%
  select(ENCODING, ITEM) %>% knitr::kable()
```

