---
title: "Word order constraints on event-internal modifiers - Experiment 1: Likert Scale study on Anaphoricity"
author: "Tibor Kiss"
date: "09.09.2021"
output:
  html_document: default
  word_document: default
---

```{r libraries, warning=FALSE, message=FALSE}
library(ordinal)
library(tidyverse)
library(emmeans)
```

#### Introduction

This document describes the necessary analysis of Experiment 1 (LS) for event-internal modifiers (second stage). We include a model with and without interaction. As can be seen below, the difference between the two models in terms of fit is just above 0.05. 


#### Read in data set

```{r data preparation}
exp_anaphoricity_ls.test.data <- 
  read.csv("../Data/LikertSkala_test.csv", fileEncoding = "UTF-8")

exp_anaphoricity_ls.test.data$CONDITION_NO <- factor(exp_anaphoricity_ls.test.data$CONDITION_NO)

exp_anaphoricity_ls.test.data$FCT_ANSWER <- 
  factor(exp_anaphoricity_ls.test.data$ANSWER, ordered = TRUE)

exp_anaphoricity_ls.test.data$ENCODING <- 
  factor(as.character(exp_anaphoricity_ls.test.data$ENCODING))

exp_anaphoricity_ls.test.data$KEY_CONDITION = relevel(exp_anaphoricity_ls.test.data$KEY_CONDITION, "PP>OBJ") ## CONDITION_NO 1

## more illustrative variable names
exp_anaphoricity_ls.test.data$POSITION <- factor(exp_anaphoricity_ls.test.data$KEY_CONDITION)
exp_anaphoricity_ls.test.data$subjects <- factor(exp_anaphoricity_ls.test.data$workerId)
exp_anaphoricity_ls.test.data$items <- factor(exp_anaphoricity_ls.test.data$ENCODING)

```

#### Empirical distribution of judgments

Number of accepted participants: `r length(unique(exp_anaphoricity_ls.test.data$workerId))`.

```{r distribution of items to participants, warning = FALSE, message = FALSE}
exp_anaphoricity_ls.dist.summary <- exp_anaphoricity_ls.test.data %>% 
  select(workerId, CONDITION_NO) %>% table() %>% as.data.frame()
exp_anaphoricity_ls.dist.summary %>% filter(Freq < 18)
```

Each participant has seen 18 test items per condition, since no participant has seen less than 18 test items per condition (8->). 

```{r distribution of judgments, fig.width = 10, message = FALSE, warning = FALSE}

exp_anaphoricity_ls.test.summary <- 
  exp_anaphoricity_ls.test.data %>% 
  group_by(POSITION, ADVERBIAL_TYPE, ANSWER) %>% 
  summarise(count = n())

ggplot(exp_anaphoricity_ls.test.summary, aes(x = ANSWER, y = count)) +
  geom_bar(stat = "identity", position = "dodge", show.legend = FALSE) +
  facet_wrap(POSITION~ADVERBIAL_TYPE) + 
  labs(x = "Ratings", y = "Frequency of Ratings")

```


### Random slope model for items and subjects without interaction

#### Definition of model

```{r model, cache = TRUE}

exp_anaphoricity_ls.test.data$ADVERBIAL_TYPE <-
  relevel(factor(exp_anaphoricity_ls.test.data$ADVERBIAL_TYPE), ref = "INSTR")
exp_anaphoricity_ls.test.data$POSITION <-
  relevel(factor(exp_anaphoricity_ls.test.data$POSITION), ref = "PP>OBJ")

exp_anaphoricity_ls.clmm <- 
  clmm(FCT_ANSWER  ~ ADVERBIAL_TYPE + POSITION + 
         (1 + POSITION| subjects) + 
         (1 + POSITION| items), 
       exp_anaphoricity_ls.test.data)

summary(exp_anaphoricity_ls.clmm)

```

#### Effects in the model w/o interaction

Since we have chosen INSTR as reference value, the values for ILOC and COM(O) both show significance, but it should be kept in mind that they are practically identical. The general effect of `POSITION == OBJ > PP` is `r round(exp(exp_anaphoricity_ls.clmm$beta[3]), 3)`. The effects for COM(O) and ILOC in `POSITION == PP > OBJ` are `r round(exp(exp_anaphoricity_ls.clmm$beta[1]), 3)`, and `r round(exp(exp_anaphoricity_ls.clmm$beta[2]), 3)`, respectively.  


#### Model predictions (without interaction)

```{r graphical presentation of model without interaction, fig.width = 10}

pred <-function(eta, theta, cat = 1:(length(theta)+1), inv.link= plogis){
  Theta <- c(-1e3, theta, 1e3)
  sapply(cat, function(j) inv.link(Theta[j+1] - eta) - inv.link(Theta[j] - eta))
}

exp_anaphoricity_ls.mat <- 
  data.frame(
    adv = rep(c(0, exp_anaphoricity_ls.clmm$beta[1], exp_anaphoricity_ls.clmm$beta[2]), 2),
    cond = c(rep(0, 3), rep(exp_anaphoricity_ls.clmm$beta[3], 3))
  )

pred.mat <- pred(eta=rowSums(exp_anaphoricity_ls.mat), 
                 theta=exp_anaphoricity_ls.clmm$Theta)

exp_anaphoricity_ls.mat$condition <- 
  c(rep("PP>OBJ", 3), rep("OBJ>PP", 3))
exp_anaphoricity_ls.mat$adv_type <- 
  rep(c("INSTR", "COM(O)", "ILOC"), 2)

exp_anaphoricity_ls.pred <- 
  cbind(exp_anaphoricity_ls.mat, pred.mat)

exp_anaphoricity_ls.pred.long <- 
  exp_anaphoricity_ls.pred %>%
  gather("1":"5", key = "ANSWER", value = "rating")

exp_anaphoricity_ls.pred.long$condition <- 
  relevel(factor(exp_anaphoricity_ls.pred.long$condition), ref = "PP>OBJ")

ggplot(exp_anaphoricity_ls.pred.long, aes(x = ANSWER, y = rating)) + 
  geom_bar(stat = "identity", show.legend = FALSE) + 
  facet_wrap(condition ~ adv_type) + 
  labs(x = "Ratings", y = "Pr(Ratings)")


```

#### Model with interaction

```{r model with interaction, cache = TRUE}

options(width = 300)

exp_anaphoricity_ls.clmm2 <- 
  clmm(FCT_ANSWER  ~ ADVERBIAL_TYPE * POSITION + 
         (1 + POSITION| subjects) + 
         (1 + POSITION| items), 
       exp_anaphoricity_ls.test.data)

summary(exp_anaphoricity_ls.clmm2)


```


#### Comparison of Models


```{r model comparison}

# Please notice that we are using anova.clm here.

comp <- anova(exp_anaphoricity_ls.clmm, exp_anaphoricity_ls.clmm2)
show(comp)
```

Analysis of deviance reveals that the model with interaction does not significantly reduce the deviance, but the value of `r round(comp$"Pr(>Chisq)"[2], 3)` is of course pretty close to the famous threshold of P = 0.05


#### Test items in disfavored positions with high ratings 

```{r}

exp_anaphoricity_ls.rating.summary <- 
  subset(exp_anaphoricity_ls.test.data, 
         ADVERBIAL_TYPE == "COM(O)" & POSITION == "PP>OBJ") %>% 
  group_by(ENCODING, ANSWER) %>% 
  summarise(count = n()) %>%
  as.data.frame()

ggplot(exp_anaphoricity_ls.rating.summary, aes(x = ANSWER, y = count)) +
  geom_bar(stat = "identity") +
  facet_wrap(~ENCODING) +
  labs(x = "Rating", y = "Frequency of rating")
```

```{r}
read.csv("../Data/stimuli_test_basic.csv", fileEncoding = "UTF-8") %>% filter(ADVERBIAL_TYPE == "COM(O)" & KEY_CONDITION == "PP>OBJ") %>% select(ENCODING, ITEM)%>% knitr::kable()
```

